---
title: "Assignment 4"
author: "Tarakeswari Ramachandra"
date: "30/11/2021"
output: html_document
---
Question 1



1 Identify all the SNPs with 'significant' associations using Bonferroni, Holm, and the Benjamini-Hochberg (BH) methods to adjust for multiple testing. Conduct inferences at 0.05 significance level (for Bonferroni's and Holm's method) and 0.05 false discovery rate for the BH method.


Report the SNPs (column numbers) that were 'significant' according to each method.



```{r}
load("/Users/tarakeswariramachandra/Downloads/Xy.RData")

n=nrow(X)
m=ncol(X)

Pvalue=rep(NA, m)
P.ADJ=Pvalue
P.ADJ.hol=Pvalue
P.ADJ.fdr=Pvalue
for(i in 1:m){
  fm=lm(y~X[,i])
  Pvalue[i]=summary(fm)$coef[2,4]
  
  

}
P.ADJ=p.adjust(Pvalue, method="bonferroni")
P.ADJ.hol=p.adjust(Pvalue, method="holm")
P.ADJ.fdr=p.adjust(Pvalue, method="fdr")
P_bon=data.frame(P.ADJ)
P_hol=data.frame(P.ADJ.hol)
P_fdr=data.frame(P.ADJ.fdr)
reject1=P_bon<0.05
reject2=P_hol<0.05
reject3=P_fdr<0.05

W=which(reject1==TRUE)
W=data.frame(W)

W1=which(reject2==TRUE)
W1=data.frame(W1)

W2=which(reject3==TRUE)
W2=data.frame(W2)

print("SNP significant through Bonferroni method")
W
print("SNP significant through Holm method")
W1
print("SNP significant through BH method")
W2
```


Question 2 : Produce the following plot (using un-adjusted p-values)

Manhattan plot (-log10(pvalue) in the y-axis versus position of the SNP (1:ncol(X)) in the horizontal axis)

A a histogram of p-values with 500 bins hist(pvalues,500)

A qqplot(), i.e., a scatterplot with the empirical quantiles of your p-values in the vertical axis versus the quantiles of the uniform distribution (i.e., the expected distribution under the null) in the x-axis. Display both axis in the -log10 scale. Add a 45-degree line.

At what value of -log10(pvalue) do you see clear departure from the null distribution?


From the QQ-plot we can see at -log10(Pvalue) between 4 and 5 points departs from normality or the 45 degree line and around the value greater than 5 it departs clearly . Since we are taking -log10 large value of (-log10) corresponds to smaller P-value. 

```{r}

par(mfrow=c(2,2))

hist(Pvalue, 500)

b=1:m
plot(y=-log10(Pvalue), b, xlab = "SNP", ylab = "-log10(Pvalue)", main = "Manhattan plot")
abline(h=-log10(max(Pvalue[P.ADJ.fdr<0.05])), lty=2)
prob=seq(from=0,to=1,by=(1/length(Pvalue)))
empQualtile=quantile(Pvalue,prob)
plot(-log10(empQualtile), x= -log10(prob), cex=0.5, col=4)
abline(a=0, b=1, col=2)
abline(h=-log10(max(Pvalue[P.ADJ.fdr<0.05])), lty=2)


#QQplot can also be produce using the cmd qqnorm

qqnorm(-log10(Pvalue))
```













