---
title: "Assignment 3"
author: "Tarakeswari Ramachandra"
date: "18/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Problem 1

Problem:1.1. Use Monte Carlo Methods (at least 5000 replicates) to estimate power curves (power versus sample size, N=30,50,100,200,500,1000,3000) by effect size (b2=0,0.2, 0.3 and 0.5 ).

```{r}
N=c(30,50,100,200,500,1000,3000)
b=c(0,0.2, 0.3, 0.5)
nRep=5000
PAR=expand.grid(N=N,b=b)
 PAR$rejRate=NA
for(s in 1:nrow(PAR)){
  n=PAR$N[s]
  b2=PAR$b[s]
  rejects=rep(NA,nRep)
  a=n/2
  for(i in 1:nRep){
    
    # Simulate data using the parameters for the s-scenario
    # Apply the decision rule
    mu=120
    b1=-3
    BMI_F=data.frame("BMI"=rnorm(a,mean=26.5,sd=sqrt(30)))
    BMI_M=data.frame("BMI"=rnorm(a,mean=27.4,sd=sqrt(16.7)))
    BMI_F$sex[1:a]=0
    BMI_M$sex[1:a]=1
    BMI=rbind(BMI_M, BMI_F)
    BMI=scale(BMI,center=TRUE,scale=FALSE)
    error=rnorm(n,mean=0, sd=sqrt(300))
    SBP=120+BMI[,2] * (-3) + (BMI[,1]-mean(BMI[,1]))* b2 + error
    fm=lm(SBP~BMI[,1]+BMI[,2])
    pVal=summary(fm)$coef[2,4]
    rejects[i]=pVal<0.05
  }
  
  # Estimate rejection rate
  PAR$rejRate[s]=mean(rejects)
  
}

plot(PAR$N, PAR$rejRate ,type="b", xlab="Sample size", ylab="Power", main="Power curve")
 
```

Problem 1.2. What is the minimum sample size needed to achieve a power of at least 80% if the effect size is 0.3?

```{r}
nRep=5000
N_min=1140
a2=N_min/2
b2=0.3
rejects2=rep(NA, nRep)
for(i in 1:nRep){
  
  # Simulate data using the parameters for the s-scenario
  # Apply the decision rule
  mu=120
  b1=-3
  BMI_F=data.frame("BMI"=rnorm(a2,mean=26.5,sd=sqrt(30)))
  BMI_M=data.frame("BMI"=rnorm(a2,mean=27.4,sd=sqrt(16.7)))
  BMI_F$sex[1:a2]=0
  BMI_M$sex[1:a2]=1
  BMI=rbind(BMI_M, BMI_F)
  BMI=scale(BMI,center=TRUE,scale=FALSE)
  error=rnorm(N_min,mean=0, sd=sqrt(300))
  SBP=120+BMI[,2] * (-3) + (BMI[,1]-mean(BMI[,1]))* b2 + error
  fm=lm(SBP~BMI[,1]+BMI[,2])
  pVal=summary(fm)$coef[2,4]
  rejects2[i]=pVal<0.05
}
Power_min=mean(rejects2)
Power_min
```


Problem 2


Assume that the effect of BMI on SBP is higher (0.4) in male than in female (0.2). Develop a MC study to estimate the power to detect this interaction as a function of sample size (go up to N=10,000). Suggestion: extend the baseline model by adding a sex by BMI interaction and assess the power to detect an interaction of the size above-specified.

Q: Can a power of at least 50% be achieved with a sample size <= 10,000?

When the sample size N was set to 10,000 I got a power of 0.7693. With this observation, I tried with N values less than 10,000 and at N=6000 I found the power to be 0.5451667. So a power of at least 50% can be achieved when sample size is at 5500. 

```{r}
nRep1=5000
N=c(5200, 5300, 5000, 5500)
power=rep(NA, length(N))
for(s in 1:length(N)){
  Ni=N[s]
  a1=Ni/2
  rejects=rep(NA, nRep1)
for(i in 1:nRep1){
  
  # Simulate data using the parameters for the s-scenario
  # Apply the decision rule
  mu=120
  b1=-3
  BMI_F=data.frame("BMI"=rnorm(a1,mean=26.5,sd=sqrt(30)))
  BMI_M=data.frame("BMI"=rnorm(a1,mean=27.4,sd=sqrt(16.7)))
  BMI_F$sex[1:a1]=0
  BMI_M$sex[1:a1]=1
  BMI=rbind(BMI_M, BMI_F)
  error=rnorm(a1,mean=0, sd=sqrt(300))
  SBP_M=120+BMI_M[,2] * (-3) +( (BMI_M[,1]-mean(BMI_M[,1]))* (0.4) )+ error
  SBP_F=120+BMI_F[,2] * (-3) + ((BMI_F[,1]-mean(BMI_F[,1]))* (0.2) )+ error
  SBP=append(SBP_M, SBP_F)
  BMI[,1]=scale(BMI[,1],center=TRUE,scale=FALSE)
  names(BMI)=c("BMI", "SEX")
  fm=lm(SBP ~ BMI[,1] + BMI[,2] + (BMI[,1] * BMI[,2]))
  pVal=summary(fm)$coef[4,4]
  rejects[i]=pVal<0.05
}
   power[s]=mean(rejects)
}

 power

```


Problem 3

In the below code I have implemented one sided test using z-statistics and P-value and obtained the respective values.


```{r}
options(warn=-1)
N=c(19000,19500,20000)
nReps=500
final=rep(NA, length(N))
for(i in 1:length(N)){
  Powers=rep(NA, nReps)
  Ni=N[i]
  for(j in 1:nReps){
    N0=0.4*Ni
    N1=0.6 * Ni
    V=rep(1, N1)
    UV=rep(0, N0)
    Vax=data.frame(append(V, UV))
    names(Vax)=c("X")
    y_1=rbinom(n=N1, size=1, prob = (0.3/500))
    y_0= rbinom(n=N0, size=1, prob = (1/500))
    y=append(y_1, y_0)
    fm=glm(y~Vax$X, family="binomial")
    pVal=summary(fm)$coef[2,4]
    Powers[j]=pVal/2 < 0.025
   
  }
  final[i]=mean(Powers)
  
}
final
N

```


```{r}
options(warn=-1)
N=c(5800,5900,6000,6200)
nReps=500
final=rep(NA, length(N))
for(i in 1:length(N)){
  Powers=rep(NA, nReps)
  Ni=N[i]
  for(j in 1:nReps){
    N0=0.4*Ni
    N1=0.6 * Ni
    V=rep(1, N1)
    UV=rep(0, N0)
    Vax=data.frame(append(V, UV))
    names(Vax)=c("X")
    y_1=rbinom(n=N1, size=1, prob = (0.3/500))
    y_0= rbinom(n=N0, size=1, prob = (1/500))
    y=append(y_1, y_0)
    fm=glm(y~Vax$X, family="binomial")
    z=summary(fm)$coef[2,3]
    Powers[j]=z > -1.96
   
  }
  final[i]=mean(Powers)
  
}
final
N
```









